---
title: "Surveillance Data Analysis"
author: "Caroline Ngina"
date: "`r Sys.Date()`"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE)
```

```{r}
pacman::p_load(readxl,lubridate,tidyverse,gt,janitor,tidytable, skimr,flextable,officer,arsenal,dplyr,tidyr, gtsummary,ggplot2, writexl)
```

```{r}
#Creating work directory
setwd("C:/Users/cnf23/Documents/SMC 2025/Data analysis/Clean data")
```

```{r}
surveillance_data <- read_excel("all_surveillance.xlsx")
```

```{r}
#unique(surveillance_data$net_avail)
```

```{r}
#surveillance_data <- surveillance_data %>% 
  #mutate(
   # age_group=case_when(child_age>=3 & child_age<12 ~"3-11",
   #   child_age>=12 &child_age<60 ~"12-59",
   #   child_age>=60 ~"above 59",
    #  TRUE~ NA_character_
   # )
 # )
```

```{r}
surveillance_data <- surveillance_data %>% 
  mutate(
    age_group=case_when(
      child_age>=3 & child_age<6 ~"3-5",
      child_age>=6 &child_age<12 ~"6-11",
      child_age>=12 &child_age<60 ~"12-59",
      child_age>=60 ~"above 59",
      TRUE~ NA_character_
    )
  )
```

```{r}
#surveillance_data <- surveillance_data %>%
#  mutate(
 #   age_group = case_when(
 #     child_age >= 3  & child_age <= 11  ~ "3-11 mo",
 #     child_age >= 12 & child_age <= 23  ~ "12-23 mo",
  #    child_age >= 24 & child_age <= 35  ~ "24-35 mo",
  #    child_age >= 36 & child_age <= 47  ~ "36-47 mo",
 #     child_age >= 48                   ~ "48+ mo",
 #     TRUE ~ NA_character_
  #  ),
   # age_group = factor(age_group, levels = c("3-11 mo", "12-23 mo", "24-35 mo", "36-47 mo", "48+ mo"))
 # )
```


```{r}
#Renaming Variables for table 1
surveillance_data <- surveillance_data %>% 
  rename(`Age(months)`=age_group,
         Gender=child_gender,
         `Has bednet for sleeping space`=net_avail,
         `Slept under a bednet`=slept_undernet,
         `Sleeping space location`=sleep_inside_outside,
         `Know a CHP in your village`=know_chp,
         `Ever heard of SMC`=heard_smc,
         `Received any SMC cycle`=receive_smc,
         `Distance to facility(minutes)`=distance_facility,
         `Respondent's gender`=respondent_gender,
         `Respondent's Age(years)`=respondent_age
         
         )
```

```{r}
#table(surveillance_data$`Respondent's gender`)
```

```{r}
#Labelling values
library(dplyr)

surveillance_data <- surveillance_data %>%
  mutate(
    dplyr::across(
      where(~ all(na.omit(.) %in% c(0, 1))),
      ~ factor(.,
               levels = c(0, 1),
               labels = c("No", "Yes"))
    )
  )
surveillance_data <- surveillance_data %>% 
  mutate(`Ever heard of SMC`=recode(as.character(`Ever heard of SMC`), "1"="Yes", "0"="No", "999"="Don't know"))

```

```{r}
#table(surveillance_data$`Respondent's gender`)
#unique(surveillance_data$`Respondent's gender`)
```

```{r}
#Changing values within a variable
         surveillance_data <- surveillance_data %>%
           mutate(`Sleeping space location`=recode(`Sleeping space location`, "Inside"="Only indoors", "Outside"="Only outdoors", "Both"="Partly both"
))

```

```{r}
#Ordering values for table1
         surveillance_data$`Has bednet for sleeping space` <- factor(
           surveillance_data$`Has bednet for sleeping space`,
           levels = c("Yes", "No"))

         surveillance_data$`Slept under a bednet` <- factor(
           surveillance_data$`Slept under a bednet`,
           levels = c("Yes", "No"))

                  surveillance_data$`Received any SMC cycle` <- factor(
           surveillance_data$`Received any SMC cycle`,
           levels = c("Yes", "No"))
                  
                    surveillance_data$`Age(months)` <- factor(
           surveillance_data$`Age(months)`,
           levels = c("3-5", "6-11", "12-59", "above 59"))
     
                  surveillance_data$`Ever heard of SMC` <- factor(
           surveillance_data$`Ever heard of SMC`,
           levels = c("Yes", "No", "Don't know"))

                 surveillance_data$`Know a CHP in your village` <- factor(
           surveillance_data$`Know a CHP in your village`,
           levels = c("Yes", "No"))
                       
                
```

```{r}
#table(surveillance_data$`Respondent's gender`)

```

## Summary table; Participant characteristics

```{r results='asis'}
tab1 <- tableby(facility_name ~  `Age(months)` + Gender +`Sleeping space location` + `Has bednet for sleeping space` +
                  `Slept under a bednet` + `Received any SMC cycle`+ `Know a CHP in your village`, data=surveillance_data)
summary(tab1, test=FALSE, pvalues = FALSE)

```

```{r}
surveillance_data <- read_excel("all_surveillance.xlsx")

```

```{r}
#Graph1:Net ownership
net_pivot <- surveillance_data %>% group_by(facility_name) %>% 
  summarise(
    pid_new=n(pid),
    net_ownership=sum(net_avail),
    net_use=sum(slept_undernet),
    net_ownership_p=round(sum(net_avail)/n(pid)*100,1),
    net_use_p=round(sum(slept_undernet)/sum(net_avail)*100,1),
    .groups = "drop"
  )

```

```{r}
#unique(surveillance_data$net_avail)

```

## Net ownership and use

### Table for net ownership and use

```{r}
gt(net_pivot)
```

### Net ownership and use graph

```{r eval=FALSE}
net_pivot_long <- net_pivot %>% select(facility_name,net_ownership_p,net_use_p) %>% 
  pivot_longer(cols = c("net_ownership_p","net_use_p"),names_to = "facility",values_to = "proportion")
ggplot(net_pivot_long,aes(x=facility_name,y=proportion,fill=facility)) + 
  geom_col(position = "dodge",color="white") + labs(x="Facility",y="Bed net ownership and Use (%)", title="") + 
  geom_text(aes(label = proportion),position = position_dodge(width = 0.8),vjust=-0.3,size=3.0, fontface = "bold") +
  scale_y_continuous(limits = c(0,110), #Define the limit of the y axis
                              breaks=seq(0,100,25),expand = c(0,0))+
  scale_fill_manual(values = c("net_ownership_p"="#00BA38","net_use_p"="#619CFF"),
                    labels=c("net_ownership_p"="Own net","net_use_p"="Used the net"),name=NULL) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 12)) +
  theme_bw()+
  theme(legend.position = "right",
                 #panel.background = element_rect(fill = "white",colour = NA),
                 panel.border = element_blank(),
                 axis.line = element_line(colour = "black"),
                 legend.text = element_text(size=10,colour = "black"),
                 legend.title =element_text(size=12,colour = "black",face="bold"),
                 axis.title.y = element_text(size = 12, color = "black", face="plain"),
                 axis.text.y = element_text(size = 10, color = "black"),
                 axis.title.x = element_text(size = 12, color = "black"),
                 axis.text.x = element_text(size = 10, color = "black", angle = 0,hjust = 0.5,vjust = 0.5, lineheight = 0.9)
                 
           )
#net_pivot_long
ggsave("Net_ownership_use.png",
  width = 11,
  height = 6,
  dpi = 300)

```


```{r}
#Grouping
grp_net <- net_pivot %>% select(facility_name, net_ownership_p,net_use_p) %>% 
 pivot_longer(cols = c("net_ownership_p","net_use_p"),names_to = "facility",values_to = "proportion") %>% 
  mutate(facility= recode(facility, net_ownership_p= "Own bed net",net_use_p="Used bed net"  ))

ggplot(grp_net,aes(x=facility_name,y=proportion,fill=facility)) + 
  geom_col(color="white") + labs(x="Facility",y="Bed net ownership and Use (%)", title="") + 
  geom_text(aes(label = proportion),vjust=-0.3,size=3.0, fontface = "bold") +
  facet_wrap(~facility, ncol=NULL, scales="free_y") +
  scale_y_continuous(limits = c(0,110), #Define the limit of the y axis
                              breaks=seq(0,100,25),expand = c(0,0))+
  scale_fill_manual(values = c("Own bed net"="#00BA38","Used bed net"="#619CFF")) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 12)) +
  theme_bw()+
  theme(legend.position = "none",
                 #panel.background = element_rect(fill = "white",colour = NA),
                 panel.border = element_blank(),
                 axis.line = element_line(colour = "black"),
                 legend.text = element_text(size=10,colour = "black"),
                 legend.title =element_text(size=12,colour = "black",face="bold"),
                 axis.title.y = element_text(size = 12, color = "black", face="plain"),
                 axis.text.y = element_text(size = 10, color = "black"),
                 axis.title.x = element_text(size = 12, color = "black"),
                 axis.text.x = element_text(size = 10, color = "black", angle = 45,hjust = 1,vjust = 1, lineheight = 0.9)
  )

ggsave("Net_ownership_use1.png",  width = 11,  height = 6,  dpi = 300)


```




```{r}
#unique(surveillance_data$slept_undernet)

```

```{r eval=FALSE}
#unique(surveillance_data$`Ever heard of SMC`)
#table(surveillance_data$`Ever heard of SMC`)
```

```{r eval=FALSE}
library(gtsummary)
library(dplyr)
tbl_summary(surveillance_data,
            include=c(`Age(months)`,Gender,`Sleeping space location`, `Has bednet for sleeping space` ,`Slept under a bednet`,`Received any SMC cycle` ),
            by="facility_name",
            #sort = all_categorical() ~ "frequency"
            ) %>% 
  bold_labels()%>% 
  modify_header(label="**Variable**") %>% 
  modify_caption(paste0("**Table 1. Child characteristics**")) %>% 
  #add_overall() %>% 
  add_n()
```



```{r  }
reasons_data <- read_excel("REASONS.xlsx",sheet=2)
```

```{r eval=TRUE}
#unique(reasons_data$rsn_missed_smc_oth)
```


```{r}
#unique(reasons_data$rsn_missed_smc_oth)
#rsn_missed_smc12","rsn_missed_smc13",
  #"rsn_missed_smc12" = "Caregiver declined",
  #"rsn_missed_smc13" = "Child was under age",
 #"rsn_missed_smc14" = "New resident",
 #"rsn_missed_smc15" = "Child on anti-TB drugs"

```

```{r eval=TRUE }
reasons_choices1 <- reasons_data %>% select(c("KEY","facility_name","rsn_missed_smc1","rsn_missed_smc2","rsn_missed_smc3","rsn_missed_smc4","rsn_missed_smc5","rsn_missed_smc6","rsn_missed_smc7","rsn_missed_smc8","rsn_missed_smc9","rsn_missed_smc10","rsn_missed_smc11","rsn_missed_smc12", "rsn_missed_smc13","rsn_missed_smc14", "rsn_missed_smc15", "rsn_missed_smc98" ))

#reasons_choices1
```

```{r eval=TRUE }

reasons_choices <- c("rsn_missed_smc1","rsn_missed_smc2","rsn_missed_smc3","rsn_missed_smc4","rsn_missed_smc5","rsn_missed_smc6","rsn_missed_smc7","rsn_missed_smc8","rsn_missed_smc9","rsn_missed_smc10","rsn_missed_smc11","rsn_missed_smc12", "rsn_missed_smc13","rsn_missed_smc14", "rsn_missed_smc15","rsn_missed_smc98")
```

```{r eval=TRUE}
reason_labels <- c(
  rsn_missed_smc1  = "Did not know about SMC",
  rsn_missed_smc2  = "CHP did not come",
  rsn_missed_smc3  = "Household head was away from home",
  rsn_missed_smc4  = "Caretaker was away from home",
  rsn_missed_smc5  = "Child was away from home",
  rsn_missed_smc6  = "Could not find fixed-point venue or too far",
  rsn_missed_smc7  = "Drugs were not available",
  rsn_missed_smc8  = "Child had side effects from previous cycle",
  rsn_missed_smc9  = "Child was ill on the day of visit",
  rsn_missed_smc10 = "Unsure about benefits of the drugs",
  rsn_missed_smc11 = "Unsure about safety of the drugs",
  rsn_missed_smc12 = "Caregiver declined",
  rsn_missed_smc13 = "Child was under age",
  rsn_missed_smc14 = "New resident",
  rsn_missed_smc15 = "Child on anti-TB drugs",
  rsn_missed_smc98 = "Other reason"
)
```

```{r eval=TRUE}
missed <- reasons_choices1 %>% pivot_longer(cols=all_of(names(reason_labels)), names_to = "Reason", values_to = "reason_value") %>% 
  mutate(Reason = recode(Reason, !!!reason_labels), Reason = factor(Reason, levels = unname(reason_labels)))  # enforce order


missed_summary <- missed %>%
  group_by(facility_name, Reason) %>%
  summarise(total = sum(reason_value, na.rm = TRUE), .groups = "drop")

  missed_wider <- missed_summary %>%
  pivot_wider(names_from = facility_name,
              values_from = total,
              values_fill = 0) %>%
  mutate(Grand_Total = rowSums(dplyr::across(where(is.numeric))))

  #missed_wider

```

## Reasons for missing SMC drug

### Reasons for missing SMC by Facility

```{r eval=TRUE}
missed_wider %>%
  gt() %>%
  tab_header(title = md("**Reasons for Missing SMC**")) %>%
  tab_options(table.font.size = 14, data_row.padding = px(6)) 

```

### Reasons for missing SMC (percentage)


```{r eval=TRUE}
v1 <- missed_wider %>% select(c("Reason", "Grand_Total")) %>% 
  filter(Reason!="Drugs were not available" ) %>% 
  mutate("Percent(%)" = round(Grand_Total/sum(Grand_Total)*100,1),
         category= case_when(
           Reason %in% c("Did not know about SMC") ~ "Awareness",
           Reason %in% c("CHP did not come", "Could not find fixed-point venue or too far") ~ "Access",
           Reason %in% c("Child had side effects from previous cycle", "Unsure about benefits of the drugs", "Unsure about safety of the drugs", "Caregiver declined") ~ "Choice",
           Reason %in% c("Child was ill on the day of visit", "Child was under age", "New resident", "Child on anti-TB drugs"  ) ~ "Eligibility",
           Reason %in% c("Caretaker was away from home", "Child was away from home", "Household head was away from home") ~ "Travel",
           Reason %in% c("Other reason") ~ "Other reason"
         )
          )  
v1 %>% #arrange(category) %>% 
  gt(rowname_col = "Reason", groupname_col = "category") %>% cols_label(Grand_Total= "Count(n)") %>%  tab_style(cell_text(weight = "bold"), locations = cells_column_labels(everything())) %>% 
  cols_align(align = "left",columns = Reason)
```


# Adherence dataset 
```{r eval=TRUE}
mainform_data <- read_excel("adherence_final_09022026.xlsx")
```

```{r eval=TRUE}
mainform_cols <- c("dateofvisit","pid","receive_smc","date_lastcycle")
mainform_subset <- mainform_data %>% select(all_of(mainform_cols)) %>% filter(receive_smc==1)
  
```

```{r eval=TRUE}
#mainform_subset <- mainform_subset %>% mutate(
# date_lastcycle=if_else(pid=="KP-SEA-1006", as.Date("2025-07-21"),date_lastcycle)
 # )
```

```{r eval=TRUE}
mainform_subset <- mainform_subset %>% mutate(
  num_days= as.numeric(difftime(dateofvisit,date_lastcycle,units = "days"))
  ) 
#%>% filter(num_days>=0)

```

```{r eval=TRUE}
date_query <- mainform_subset %>% filter(num_days<0)

```

```{r eval=TRUE}
pivot_days <- mainform_subset %>% group_by(num_days) %>%
  summarise(num_children=n_distinct(pid), .groups= "drop")
```

## How long after SMC do children come back

### Number of Days After SMC

```{r eval=TRUE}
ggplot(pivot_days, aes(x=num_days, y=num_children))+ 
  theme_bw() +
  geom_col(fill="#619CFF")+
             labs(x="Days since last SMC", y="Number of children", title="")+#blank fill removes the legend title
           scale_y_continuous(limits = c(0,20), #Define the limit of the y axis
                              breaks=seq(0,30,2),expand = c(0,0))+
  scale_x_continuous(limits = c(0,150), #Define the limit of the y axis
                              breaks=seq(0,150,7),expand = c(0,0))+
           theme(legend.position = "right",
                 panel.border = element_blank(),
                 axis.line = element_line(colour = "black"),
                 legend.text = element_text(size=10,colour = "black"),
                 legend.title =element_text(size=12,colour = "black",face="bold"),
                 axis.title.y = element_text(size = 12, color = "black"),
                 axis.text.y = element_text(size = 10, color = "black"),
                 axis.title.x = element_text(size = 12, color = "black"),
                 axis.text.x = element_text(size = 10, color = "black", angle = 0,margin =margin(t=2,r=8,b=2,l=4,"pt"))
                 
           )



```

```{r eval=FALSE}
#pivot_days <- pivot_days %>% mutate(
 # Days = case_when(
 #   num_days >=0 & num_days <=14 ~ "0-14",
 #   num_days >=15 & num_days <=29 ~ "15-29",
  #  num_days >=30 & num_days <=44 ~ "30-44",
  #  num_days >=45 & num_days <=59 ~ "45-59",
  #  num_days >=60 & num_days <=74 ~ "60-74",
  #  num_days >=75 & num_days <=89 ~ "75-89",
   # num_days >=90  ~ "90+",
#  )
#)
```

```{r}
pivot_days <- pivot_days %>%
  mutate(
    Days = case_when(
      num_days >= 0   & num_days <= 7   ~ "0-7",
      num_days >= 8   & num_days <= 15  ~ "8-15",
      num_days >= 16  & num_days <= 23  ~ "16-23",
      num_days >= 24  & num_days <= 31  ~ "24-31",
      num_days >= 32  & num_days <= 39  ~ "32-39",
      num_days >= 40  & num_days <= 47  ~ "40-47",
      num_days >= 48  & num_days <= 55  ~ "48-55",
      num_days >= 56  & num_days <= 63  ~ "56-63",
      num_days >= 64  & num_days <= 71  ~ "64-71",
      num_days >= 72  & num_days <= 79  ~ "72-79",
      num_days >= 80  & num_days <= 87  ~ "80-87",
      num_days >= 88  ~ "88+",
      TRUE ~ NA_character_
    )
  )

```


```{r}
days_grouped <- pivot_days %>%  group_by(Days) %>% summarise(Children=sum(num_children)) %>% 
  mutate(
    Days = factor(
      Days,
      levels = c(
        "0-7","8-15","16-23","24-31","32-39","40-47","48-55",
        "56-63","64-71","72-79","80-87","88+"
      ),
      ordered = TRUE
    )
  )

```

### Number of Days After SMC (grouped)

```{r}
ggplot(days_grouped, aes(x=Days, y=Children))+ 
  theme_bw() +
  geom_col(fill="#619CFF")+
    geom_text(aes(label = Children),vjust=-0.5,size=3.0, fontface = "bold") +

             labs(x="Days since last SMC", y="Number of children", title="")+#blank fill removes the legend title
           scale_y_continuous(limits = c(0,120), #Define the limit of the y axis
                              breaks=seq(0,120,20),expand = c(0,0))+
           theme(legend.position = "right",
                 panel.border = element_blank(),
                 panel.grid = element_blank(),
                 axis.line = element_line(colour = "black"),
                 legend.text = element_text(size=10,colour = "black"),
                 legend.title =element_text(size=12,colour = "black",face="bold"),
                 axis.title.y = element_text(size = 12, color = "black"),
                 axis.text.y = element_text(size = 10, color = "black"),
                 axis.title.x = element_text(size = 12, color = "black"),
                 axis.text.x = element_text(size = 10, color = "black", angle = 0,margin =margin(t=2,r=8,b=2,l=4,"pt"))
                 
           )

```

```{r}
adherence_data <- read_excel("adherence_final_09022026.xlsx")
```


```{r eval=TRUE}
#table(surveillance_data$give_remain_tablets)
adherence <- xtabs(~give_remain_tablets, adherence_data)         
 #adherence        
  adherence <- as.data.frame(adherence) 
  
  adherence_p <- adherence %>% 
    filter(!is.na(Freq)) %>% 
    mutate(percent=round(Freq/sum(Freq)*100,2),
           give_remain_tablets=recode(give_remain_tablets, "1"="Yes, all", "2"="Yes, some", "3"="No, none"))
  #adherence_p
  
  adherence_p$give_remain_tablets <- factor(
    adherence_p$give_remain_tablets,
    levels = c("Yes, all", "Yes, some", "No, none"))        

```

## SMC Drug Adherence

### Table

```{r}
gt(adherence_p)
```

## SMC drug adherence chart

```{r}
ggplot(adherence_p, aes(x=give_remain_tablets, y=percent, fill=give_remain_tablets))+
  
    geom_bar(stat = "identity", width = 1, color = "white")+
    geom_text( aes(label = paste0(percent, "%")), vjust = -0.5, size = 3.0, fontface = "bold" )+

    theme_bw()+
  
    labs(x="Took remaining doses", y="Children (%)", fill="", title="")+#blank fill removes the legend title
  
    scale_y_continuous(limits = c(0,100), #Define the limit of the y axis
                       breaks=seq(0,100,20),expand = c(0,0))+

    theme(legend.position = "right",
          panel.border = element_blank(),
          axis.line = element_line(colour = "black"),
          legend.text = element_text(size=10,colour = "black"),
          legend.title =element_text(size=12,colour = "black",face="bold"),
          axis.title.y = element_text(size = 12, color = "black"),
          axis.text.y = element_text(size = 10, color = "black"),
          axis.title.x = element_text(size = 12, color = "black"),
          axis.text.x = element_text(size = 10, color = "black", angle = 0,margin =margin(t=2,r=8,b=2,l=4,"pt"))
    )
```

## SMC coverage by month

```{r}
surveillance_data <- read_excel("all_surveillance.xlsx")
```

```{r}
#table(surveillance_data$receive_smc)
###Generating visit month
surveillance_data <- surveillance_data %>%
  mutate(
    dateofvisit1 = ymd(dateofvisit),  # make sure it's a proper date
    month_visit = month(dateofvisit, label = TRUE, abbr = FALSE),
    SMC_cycle = case_when(
      month(dateofvisit) == 6  ~ "Cycle 1",
      month(dateofvisit) == 7  ~ "Cycle 2",
      month(dateofvisit) == 8  ~ "Cycle 3",
      month(dateofvisit) == 9  ~ "Cycle 4",
      month(dateofvisit) == 10 ~ "Cycle 5",
      TRUE ~ "Outside SMC period"
    )
  )
```

```{r }
#to find out if all are within the smc cycle
#surveillance_data %>%
  #summarise(
  #  total_records = n(),
  #  missing_SMC = sum(is.na(SMC_cycle) | SMC_cycle == "Outside SMC period"),
   # with_SMC = sum(!is.na(SMC_cycle) & SMC_cycle != "Outside SMC period")
 # )


```

```{r}
#if there are missing smc cycle look them up
#missing_SMC <- surveillance_data %>%
#  filter(is.na(SMC_cycle) | SMC_cycle == "Outside SMC period") %>%
 # select(facility_name, id_enter_type, id_code_enter, id_code_scan)

#missing_SMC

```

```{r}
surveillance_data <- surveillance_data %>%
  mutate(
    month_visit = month(dateofvisit, label = TRUE, abbr = FALSE),

    SMC_status = case_when(
      # June (expected 1 dose) ---
      month_visit == "June" & smc_cycles >= 1 ~ "Received all",
      month_visit == "June" & (is.na(smc_cycles) | smc_cycles == "") ~ "Missed all",

      # July (expected 2 doses) ---
      month_visit == "July" & smc_cycles >= 2 ~ "Received all",
      month_visit == "July" & smc_cycles == 1 ~ "Received some",
      month_visit == "July" & (is.na(smc_cycles) | smc_cycles == "") ~ "Missed all",

      #  August (expected 3 doses) ---
      month_visit == "August" & smc_cycles >= 3 ~ "Received all",
      month_visit == "August" & smc_cycles %in% 1:2 ~ "Received some",
      month_visit == "August" & (is.na(smc_cycles) | smc_cycles == "") ~ "Missed all",

      # September (expected 4 doses) ---
      month_visit == "September" & smc_cycles >= 4 ~ "Received all",
      month_visit == "September" & smc_cycles %in% 1:3 ~ "Received some",
      month_visit == "September" & (is.na(smc_cycles) | smc_cycles == "") ~ "Missed all",

      #  October (expected 5 doses) ---
      month_visit == "October" & smc_cycles >= 5 ~ "Received all",
      month_visit == "October" & smc_cycles %in% 1:4 ~ "Received some",
      month_visit == "October" & (is.na(smc_cycles) | smc_cycles == "") ~ "Missed all",

      TRUE ~ NA_character_
    )
  )

```

```{r}
smc_summary <- surveillance_data %>%
  group_by(month_visit, SMC_status) %>%
  summarise(n = n(), .groups = "drop") %>%
  pivot_wider(
    names_from = SMC_status,
    values_from = n,
    values_fill = 0
  ) %>%
  rename(
    "Month" = month_visit,
    "Missed all" = "Missed all",
    "Received some" = "Received some",
    "Received all" = "Received all"
  )

#smc_summary

```

```{r}
smc_summary <- surveillance_data %>%
  mutate(
    month_visit = month(dateofvisit, label = TRUE, abbr = FALSE),
    month_visit = if_else(month_visit == "June", "July", month_visit)   # <-- Combine June into July
  ) %>%
  filter(month(dateofvisit1) >= 6 & month(dateofvisit1) <= 10) %>% 
  group_by(month_visit, SMC_status) %>%
  summarise(n = n(), .groups = "drop_last") %>%
  mutate(perc = 100 * n / sum(n)) %>%
  ungroup() %>%
  pivot_wider(
    names_from = SMC_status,
    values_from = c(n, perc),
    values_fill = 0,
    names_glue = "{.value}_{gsub(' ', '_', SMC_status)}"
  ) %>%
  mutate(dplyr::across(starts_with("perc_"), round, 1)) %>%
  rowwise() %>%
  mutate(
    total = sum(dplyr::across(starts_with("n_")))
  ) %>%
  ungroup() %>%
  transmute(
    Month = factor(month_visit,
                   levels = c("July", "August", "September", "October"),
                   ordered = TRUE),
    `Missed all` = paste0(n_Missed_all, " (", perc_Missed_all, "%)"),
    `Received some` = paste0(n_Received_some, " (", perc_Received_some, "%)"),
    `Received all` = paste0(n_Received_all, " (", perc_Received_all, "%)"),
    `Total` = total
  ) %>%
  arrange(Month)
```

```{r }
#data for plotting
smc_plot_data <- surveillance_data %>%
  mutate(
    month_visit = month(dateofvisit, label = TRUE, abbr = FALSE),
    month_visit = if_else(month_visit == "June", "July", month_visit)
  ) %>%
  filter(month(dateofvisit1) >= 6 & month(dateofvisit1) <= 10) %>%
  group_by(month_visit, SMC_status) %>%
  summarise(n = n(), .groups = "drop") %>%
  ungroup()
```

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

# ---- PREPARE DATA ----
smc_plot_long <- smc_plot_data %>%
  mutate(
    SMC_status = case_when(
      SMC_status == "Missed all" ~ "Missed all",
      SMC_status == "Received some" ~ "Received some",
      SMC_status == "Received all" ~ "Received all",
      TRUE ~ SMC_status
    )
  ) %>%
  
  # spread wide
  pivot_wider(
    names_from = SMC_status,
    values_from = n,
    values_fill = 0
  ) %>%
  
  # month order
  mutate(
    Month = factor(
      month_visit,
      levels = c("July", "August", "September", "October"),
      ordered = TRUE
    )
  ) %>%
  
  # return to long
  pivot_longer(
    cols = c(`Missed all`, `Received some`, `Received all`),
    names_to = "SMC_status",
    values_to = "count"
  ) %>%
  
  # REAL stacking order here
  mutate(
    SMC_status = factor(
      SMC_status,
      levels = c("Missed all","Received some", "Received all" ),
      ordered = TRUE
    ))
```

```{r}
# ---- PLOT ----
plot_smc_month <- ggplot(smc_plot_long, aes(x = Month, y = count, fill = SMC_status)) +
  geom_col(width = 0.7, color = "white") +
  theme_bw()+
  labs(x = "Month", y = "Number of Children", fill = "SMC Status", title = "") +
  scale_y_continuous(limits = c(0,300), #Define the limit of the y axis
                              breaks=seq(0,300,50),expand = c(0,0))+
  scale_fill_manual(values = c("#f8766d","#00BA38","#619CFF"), labels=c("Missed all","Received some", "Received all"),name=NULL) +

  theme_minimal(base_size = 16) +
  theme(
     axis.line = element_line(colour = "black"),
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.text = element_text(color = "black")
  )

# ---- SAVE HIGH RESOLUTION ----
ggsave(
  filename = "smc_coverage_by_month.png",
  plot = plot_smc_month,
  width = 9,
  height = 6,
  dpi = 600
)
plot_smc_month
```



## Pie chart
```{r evail=FALSE}
adherence_data <- read_excel("adherence_final_09022026.xlsx")
```

```{r eval=FALSE}
#table(surveillance_data$receive_smc)
###Generating visit month
adherence_data <- adherence_data %>%
  mutate(
    #dateofvisit1 = ymd(dateofvisit),  # make sure it's a proper date
    month_visit = month(dateofvisit, label = TRUE, abbr = FALSE),
    SMC_cycle = case_when(
      month(dateofvisit) == 6  ~ "Cycle 1",
      month(dateofvisit) == 7  ~ "Cycle 2",
      month(dateofvisit) == 8  ~ "Cycle 3",
      month(dateofvisit) == 9  ~ "Cycle 4",
      month(dateofvisit) == 10 ~ "Cycle 5",
      TRUE ~ "Outside SMC period"
    )
  )
```

```{r}
#to find out if all are within the smc cycle
#surveillance_data %>%
  #summarise(
  #  total_records = n(),
  #  missing_SMC = sum(is.na(SMC_cycle) | SMC_cycle == "Outside SMC period"),
   # with_SMC = sum(!is.na(SMC_cycle) & SMC_cycle != "Outside SMC period")
 # )
#unique(adherence_data$SMC_cycle)
#unique(adherence_data$month_visit)
```

```{r}
#if there are missing smc cycle look them up
#missing_SMC <- surveillance_data %>%
#  filter(is.na(SMC_cycle) | SMC_cycle == "Outside SMC period") %>%
 # select(facility_name, id_enter_type, id_code_enter, id_code_scan)

#missing_SMC
```

```{r eval=FALSE}
adherence_data <- adherence_data %>%
  mutate(
    SMC_status = case_when(
      # June (expected 1 dose) ---
      month_visit == "June" & smc_cycles >= 1 ~ "Received all",
      month_visit == "June" & (is.na(smc_cycles) | smc_cycles == "") ~ "Missed all",

      # July (expected 2 doses) ---
      month_visit == "July" & smc_cycles >= 2 ~ "Received all",
      month_visit == "July" & smc_cycles == 1 ~ "Received some",
      month_visit == "July" & (is.na(smc_cycles) | smc_cycles == "") ~ "Missed all",

      #  August (expected 3 doses) ---
      month_visit == "August" & smc_cycles >= 3 ~ "Received all",
      month_visit == "August" & smc_cycles %in% 1:2 ~ "Received some",
      month_visit == "August" & (is.na(smc_cycles) | smc_cycles == "") ~ "Missed all",

      # September (expected 4 doses) ---
      month_visit == "September" & smc_cycles >= 4 ~ "Received all",
      month_visit == "September" & smc_cycles %in% 1:3 ~ "Received some",
      month_visit == "September" & (is.na(smc_cycles) | smc_cycles == "") ~ "Missed all",

      #  October (expected 5 doses) ---
      month_visit == "October" & smc_cycles >= 5 ~ "Received all",
      month_visit == "October" & smc_cycles %in% 1:4 ~ "Received some",
      month_visit == "October" & (is.na(smc_cycles) | smc_cycles == "") ~ "Missed all",

      TRUE ~ NA_character_
    )
  )
```

```{r}
#unique(adherence_data$SMC_status)
#n_a <- adherence_data %>% filter(is.na(SMC_status))
```

```{r}
#mainform_cols <- c("dateofvisit","pid","receive_smc","date_lastcycle")
#mainform_subset <- mainform_data %>% select(all_of(mainform_cols)) %>% filter(receive_smc==1)
```

```{r eval=FALSE}
#adherence_data <- adherence_data %>% mutate(
# date_lastcycle=if_else(pid=="KP-SEA-1006", as.Date("2025-07-21"),date_lastcycle)
 # )

```


```{r}
adherence_data <- adherence_data %>% mutate(
  num_days= as.numeric(difftime(dateofvisit,date_lastcycle,units = "days"))
  ) 

```


```{r eval=FALSE}
write_xlsx(x=adherence_data, "df_adherence_data.xlsx")
```


